---
title: "20210323-p8158_hw6_jsg2145"
author: "Jared Garfinkel"
date: "3/23/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(psych)
library(lavaan)
library(semPlot)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# read in data

```{r}
data = read_csv("./data/SEM for uwcb.csv") %>% 
  janitor::clean_names()
```

```{r}
skimr::skim(data)
head(data)
```

```{r}
df = data %>% 
  mutate(across(.fns = as.numeric))
```

```{r}
names(df)
```


# SEM

```{r notes, include = FALSE, eval = FALSE}
model <- ' 
  # latent variable definitions
     outuwcb =~ diet_yr + uwcb_yr
     wtteas =~ teasewt + fam_wt + bmi_sf
     bodysat =~ bmi_sf
     outuwcb ~ bodysat + wtteas
     bodysat ~ wtteas
     bmi_sf ~~ wtteas
'

#      wtteas ~ outuwcb + bodysat
# '
#      bmi_sf ~ outuwcb + bodysat + wtteas
#      
#      bodysat ~ outuwcb
#      wtteas ~ teasewt + fam_wt + outuwcb + bodysat + bmi_sf
# '

model <- ' 
     outuwcb =~ bmi_sf
     bodysat =~ bmi_sf
     bodysat ~ wtteas
     wtteas =~ bmi_sf
     outuwcb =~ bmi_sf
     outuwcb ~ bodysat + wtteas
     diet_yr 
     wtteas =~ teasewt + fam_wt + bmi_sf 
     bodysat ~ outuwcb
     wtteas ~ outuwcb + bodysat
     bmi_sf ~~ wtteas + bodysat + outuwcb
'
```

```{r}
model1 <- '
  # measurement model
    outuwcb =~ diet_yr + uwcb_yr
    wtteas =~ teasewt + fam_wt

  # correlating the exogenous variables
    wtteas ~~ bmi_sf

  # structural model - direct effects
    outuwcb ~ a*wtteas + b*bodydiss + c*bmi_sf
    bodydiss ~ d*bmi_sf + e*wtteas

  #indirect
    indirect_outuwcb := b * (d+e)
    
  #total
    total_outuwcb := a+c+b*(d+e)
'
```

```{r}
model.fit <- sem(model1, data = df, sample.cov=TRUE, missing = "ML")
```

```{r}
summary(model.fit, standardized = TRUE)
```

# Part 2, Question a

The standardized direct effects on outuwcb are 0.170 for wtteas, -0.031 for bodydiss and 0.015 for bmi_sf.

The standardized indirect effects on outuwcb are -0.031 * -0.336  = 0.010 for bmi_sf and -0.031 * -4.986 = 0.155 for wtteas.

So, the total effects on outuwcb is:

0.17 + -0.031 + 0.015 + 0.010 + 0.155 = .319

```{r}
semPaths(model.fit)
```

# Part 2, Question b

```{r}
model2 <- '
  # measurement model
    outuwcb =~ diet_yr + uwcb_yr
    wtteas =~ teasewt + fam_wt

  # structural model - direct effects
    outuwcb ~ a*wtteas + b*bodydiss + c*bmi_sf
    bodydiss ~ d*bmi_sf + e*wtteas

  #indirect
    indirect_outuwcb := b * (d+e)
    
  #total
    total_outuwcb := a+c+b*(d+e)
'
```

```{r}
model.fit2 = sem(model2, data = df, sample.cov=TRUE, missing = "ML")
```

```{r}
summary(model.fit2, standardized = TRUE)
```

The standardized direct effects are 0.163 for wtteas, -0.031 for bodysat and -0.020 for bmi_sf.

The standardized indirect effects are -0.031 * -0.465 = 0.014 for bmi_sf and -0.031 * -4.756 = 0.147 for wtteas.

So, the total effects on outuwcb are:

0.163+-0.031+0.020+0.014+0.147 = .313

These estimates of total effects are quite close (0.313 vs. 0.319)

The fit statistics are higher than the first model (Chi-squared statistic = 253 > 17.4).

```{r}
semPaths(model.fit2)
```

# Question 2, part c

```{r, include = FALSE, eval = FALSE}
model3 <- '
  # structural model - direct effects
    uwcb_yr ~ a*bodydiss + b*teasewt + c*bmi_sf
    diet_yr ~ d*bodydiss+ e*teasewt + f*bmi_sf
    bodydiss ~ g*bmi_sf + h*teasewt
    teasewt ~ j*bmi_sf
    
  # correlating the exogenous variables
    uwcb_yr ~~ diet_yr
    
  # indirect effects
    ind_uwcb_yr := j*b
    ind_diet_yr := j*e
    ind_bodydiss := h*j
    
  # total effects
    tot_uwcbyr := c+a+j*b
    tot_dietyr := f+d+j*e
    tot_bodydiss := g+h*j

'
```


```{r}
model3 <- '
  # structural model - direct effects
    uwcb_yr ~ a*bodydiss + b*teasewt + c*bmi_sf
    diet_yr ~ d*bodydiss+ e*teasewt + f*bmi_sf
    bodydiss ~ g*bmi_sf + h*teasewt
    teasewt ~ j*bmi_sf

  # indirect effects
    ind_uwcb_yr := a*g + a*h + b*j
    ind_diet_yr := d*g + d*h + e*j

  #total effects
    tot_uwcbyr := b*j+c+a*(g+h)
    tot_dietyr := e*j+f+d*(g+h)
    
  # correlating the exogenous variables
    uwcb_yr ~~ diet_yr
'
```

```{r}
model.fit3 = lavaan::sem(model3, data = df, sample.cov=TRUE, missing = "ML")
```

```{r}
summary(model.fit3, standardized = TRUE)
```

The direct effects for uwcb_yr are -0.055 for bodydiss, 0.107 for teaswt, and 0.024 for bmi_sf. 

The indirect effects for uwcb_yr are $-0.055*-0.491 + 0.107*0.054$ = `r -0.055*-0.491 + 0.107*0.054` for bmi_sf 

and, $-0.055*-2.028$ = `r -0.055*-2.028` for teasewt

The total effects are therefore -0.055 + 0.107 + 0.024 + `r -0.055*-0.491 + 0.107*0.054` + `r -0.055*-2.028` = 0.221 for uwcb_yr.

The direct effects for diet_yr are -0.034 for bodydiss, 0.033 for teaswt, and 0.024 for bmi_sf. 

The indirect effects for diet_yr are $-0.034*-0.491 + 0.033*0.054= 0.018$ for bmi_sf 

and, $-0.034*-2.028 = 0.069$ for teasewt

The total effects are therefore -0.034 + 0.033 + 0.024 + 0.018 + 0.069 = .110 for diet_yr.

```{r}
semPaths(model.fit3)
```

```{r}
fitmeasures(model.fit3, "df")
fitmeasures(model.fit3, "chisq")
```


This test is different than that in part b since there are no latent variables. There are no degrees of freedom and no test statistics. It is essentially a linear regression.